---
title: "AAEC 6950 HW2 Q3"
author: "wunderwood"
date: "2024-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Set wd, libraries
```{r}
setwd("C:\\Users\\Will Underwood\\OneDrive\\Documents\\AAEC 6960\\HW2 Q3")
library(dplyr)
library(tidyverse)
library(tidyr)
```

# Data Cleaning

### Import Data
```{r}
cereals0508 = read_csv("TradeData_9_24_2024_9_17_39.csv")
cereals0920 = read_csv("TradeData_9_24_2024_9_19_36.csv")
gdpPHW = read_csv("productivity.csv")
laborForce = read_csv("laborForce.csv")
capitalFM = read_csv("capitalFormation.csv", na = "..")
elevUnder5 = read_csv("elevationData.csv", skip = 4)
```

### get full cereal data
```{r}
cerials0520 = rbind(cereals0508, cereals0920)
```

### rename, select columns for leftjoin
```{r}
gdpPHW = gdpPHW %>% rename(period = TIME_PERIOD, reporterISO = LOCATION, gdpPerHr = OBS_VALUE)
gdpPHW = gdpPHW %>% select(c("period","reporterISO","gdpPerHr"))
```

### join gdpPHW to cereal data
```{r}
intrCer1 = cerials0520 %>% left_join(gdpPHW, by = c("reporterISO"="reporterISO", "period" = "period"))
intrCer3 = intrCer1 %>% left_join(gdpPHW, by = c("partnerISO"="reporterISO", "period" = "period"))
intrCer3 = intrCer3 %>% rename(PgdpPH = gdpPerHr.y)
intrCer3 = intrCer3 %>% rename(gdpPerHr = gdpPerHr.x)
```

### prepare capitalFM for pivot
```{r}
capitalFM = select(capitalFM, !c('Series Name', 'Series Code', 'Country Name'))
capitalFM = capitalFM %>% rename(reporterISO =`Country Code`)
```

### Pivot
```{r}
capitalFMlong = capitalFM %>% 
  pivot_longer(
    !reporterISO,
    names_to = "period",
    values_to = "capForm"
  )
```

### make year values numeric
```{r}
capitalFMlong$period = as.double(capitalFMlong$period)
```

### check for and remove duplicate rows in capitalFMlong
```{r}
capitalFMlong %>%
  count(reporterISO, period) %>% 
  filter(n>1)

capitalFMlong = capitalFMlong %>% 
  group_by(reporterISO, period) %>% 
  summarise(capForm = first(capForm), .groups = 'drop')
```

### Merge intermediate cereal 3 with capitalFMlong
```{r}
intrCer2 = intrCer3 %>% left_join(capitalFMlong, by = c("reporterISO","period"))
intrCer4 = intrCer2 %>%  left_join(capitalFMlong, by = c("partnerISO"= "reporterISO", "period" = "period"))
intrCer4 = intrCer4 %>% rename(capForm = capForm.x)
intrCer4 = intrCer4 %>% rename(PcapForm = capForm.y)

```

### prepare laborForce for pivot
```{r}
laborForce = laborForce %>% select(!c("Series Name","Series Code", "Country Name"))
laborForce = laborForce %>% rename("reporterISO" = "Country Code")

```

### pivot
```{r}
laborForceLong = laborForce %>% 
  pivot_longer(
    !reporterISO,
    names_to = "period",
    values_to = "labForce"
  )
```

### make year numeric
```{r}
laborForceLong$period = as.numeric(laborForceLong$period)
```


### merge intermediate cereal 4 with laborForce
```{r}
intrCer5 = intrCer4 %>% left_join(laborForceLong, by = c("reporterISO", "period"))
q3data = intrCer5 %>% left_join(laborForceLong, by = c("partnerISO"= "reporterISO", "period"))
q3data = q3data %>% rename(labForce = labForce.x)
q3data = q3data %>% rename(PlabForce = labForce.y)
q3data = q3data %>% rename(exports = primaryValue)
```

### create tech ratio & capital intensity ratio vars, log(those) & log(exports)
```{r}
q3data = q3data %>% mutate(techRatio = gdpPerHr/PgdpPH, 
                           capInt = capForm/labForce,
                           PcapInt = PcapForm/PlabForce,
                           capIntRatio = capInt/PcapInt,
                           LtechRatio = log(techRatio),
                           LcapIntRatio = log(capIntRatio),
                           Lexports = log(exports))
```

### elevation cleaning
```{r}
elevUnder5 = elevUnder5 %>% select(c("Country Code", "2015"))

elevUnder5 = elevUnder5 %>% rename("elevU5" = "2015")

```
```{r}
q3data = q3data %>% left_join(elevUnder5, by = c("reporterISO" = "Country Code"))
q3data = q3data %>% left_join(elevUnder5, by = c("partnerISO" = "Country Code"))

q3data = q3data %>% rename("elevU5"= "elevU5.x", "PelevU5" = "elevU5.y")
```

### remove na & self-trading
```{r}
q3data = na.omit(q3data) #omitted rows: partner country is non-oecd, thus not in labor, capital, or tech data
q3data = q3data %>% filter(reporterISO != partnerISO)
```

# Q3.a.
### i.
### summarize
```{r}
q3dataX = filter(q3data, flowCode == "X")
summary(q3dataX)
```

### visualize
```{r}
ggplot(data= q3dataX) +
  geom_point(aes(techRatio, exports, color = factor(period)))
ggplot(data = q3dataX)+
  geom_point(aes(capIntRatio, exports, color = factor(period)))
```

# ii
```{r}
Loglog = lm(Lexports ~ LtechRatio + LcapIntRatio, data = q3dataX)
Linlog = lm(exports ~ LtechRatio + LcapIntRatio, data = q3dataX)
Linlin = lm(exports ~ techRatio + capIntRatio, data = q3dataX)

summary(Loglog)
summary(Linlog)
summary(Linlin)
```
# Q3.d
```{r}
Linlin2 = lm(exports ~ techRatio + capIntRatio + elevU5 , data = q3dataX)

summary(Linlin2)
```





